---
title: "gtsummary in quarto"
author: "Charlotte Doran"
format:
  html:
    theme: cosmo
execute: 
  echo: false
  message: false
  warning: false
  error: false
  cache: true
---

```{r, setup}
library(tidyverse)
library(gtsummary)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd", 
               "id", "nsibs", "samp", "race_eth", "sex", "region", 
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"), 
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |> 
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes"))) 

table1 <- tbl_summary(
	#nlsy was capitalized and threw and error
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))

stats <- list(n = nrow(nlsy), med_income = median(nlsy$income, na.rm = TRUE))
```

# Customization of `tbl_summary()`

This study used data from the National Longitudinal Survey of Youth, which contains data from `r format(stats$n, big.mark = ",")` participants with a median income of \$`r format(round(stats$med_income, 0), big.mark = ",")`. An overview of the study population can be found in @tbl-1. Over half of male participants (`r gtsummary::inline_text(table1, variable = "race_eth_cat", column = "Male", level = "Non-Black, Non-Hispanic", pattern = "{n} [{p}%]")`) were non-Black, non-Hispanic.

```{r}
#| label: tbl-1
#| tbl-cap: Demographic characteristics of NLSY participants

table1
```

**Table 2**

```{r, table2}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

**Table 3**

```{r, table3}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing") |> 
  add_p(test = list(all_continuous() ~ "t.test", 
                    all_categorical() ~ "chisq.test")) |> 
  add_overall(col_label = "**Total**") |> 
  bold_labels() |> 
  modify_footnote(update = everything() ~ NA) |> 
  modify_header(label = "**Variable**", p.value = "**P**")
```

# Univariate regression

**Table 4**

```{r, table4}
tbl_uvregression(
  nlsy, 
  y = income,
  include = c(sex_cat, race_eth_cat,
              eyesight_cat, income, age_bir),
  method = lm)
```

**Table 5**

```{r, table5}
tbl_uvregression(
  nlsy, 
  y = glasses,
  include = c(sex_cat, race_eth_cat,
              eyesight_cat, glasses, age_bir),
  method = glm,
  method.args = list(family = binomial()),
  exponentiate = TRUE)
```

## Multivariable regressions

## Some regressions

```{r, linear}
linear_model <- lm(income ~ sex_cat + age_bir + race_eth_cat, 
                   data = nlsy)
```

```{r, linear_interaction}
linear_model_int <- lm(income ~ sex_cat*age_bir + race_eth_cat, 
                   data = nlsy)
```

```{r, logistic}
logistic_model <- glm(glasses ~ eyesight_cat + sex_cat + income, 
                      data = nlsy, family = binomial())
```

## Tables

**Table 6**

```{r, table6}
tbl_regression(
  linear_model, 
  intercept = TRUE,
  label = list(
    sex_cat ~ "Sex",
    race_eth_cat ~ "Race/ethnicity",
    age_bir ~ "Age at first birth"
  ))
```

**Table 7**

```{r, table7}
tbl_regression(
  logistic_model, 
  exponentiate = TRUE,
  label = list(
    sex_cat ~ "Sex",
    eyesight_cat ~ "Eyesight",
    income ~ "Income"
  ))
```

**Table 8**

```{r, table8_prep}
tbl_no_int <- tbl_regression(
  linear_model, 
  intercept = TRUE,
  label = list(
    sex_cat ~ "Sex",
    race_eth_cat ~ "Race/ethnicity",
    age_bir ~ "Age at first birth"
  ))

tbl_int <- tbl_regression(
  linear_model_int, 
  intercept = TRUE,
  label = list(
    sex_cat ~ "Sex",
    race_eth_cat ~ "Race/ethnicity",
    age_bir ~ "Age at first birth",
    `sex_cat:age_bir` ~ "Sex/age interaction"
  ))
```

```{r, table8}
tbl_merge(list(tbl_no_int, tbl_int), 
          tab_spanner = c("**Model 1**", "**Model 2**"))
```
